Nail: A practical interface generator for data formats

### 2014-09-10 paper/osdi14/intro.tex
abstr and intro nits

Code that handles untrusted inputs, such as processing network
data or parsing a file, is error-prone and is often exploited by
attackers.
%   This is in part because attackers have precise control
% over the inputs to that code, and can craft inputs that trigger
% subtle corner cases in input processing.
For example, the libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013~\cite{cvedetails:libpng},
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past~\cite{cvedetails:zlib}.
With a memory-unsafe language like C, mistakes in input processing
code can lead to memory errors like buffer overflows, but even with
a memory-safe language like Java, inconsistencies between different
parsers can lead to security issues~\cite{saurik-masterkey}.

A promising approach to avoid these problems is to specify
a precise grammar for the input data format, and to use a parser
generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
processing code.  Developers that use a parser generator do not need
to write error-prone input processing code on their own, and as long
as the parser generator is bug-free, the application will be safe from
input processing vulnerabilities. Grammars can be re-used between
applications, further reducing effort and eliminating ambiguities.

@@

Second,  applications often need to produce output in the same
format as their input---for example, applications might both
read and write files, or both receive and send network packets.
Most parser generators focus on just parsing an input, rather
than producing an output, thus requiring the programmer to manually
construct outputs, which is work-intensive and more code that could contain errors.
Some parser generators, such as
Boost.Spirit~\cite{boost-spirit},
allow reusing the grammar for generating output from the internal
representation.  However, those generators require yet another set of
semantic actions to be written, transforming the internal representation
back into an AST\@.

Third, many data formats contain redundancies, such as repeating
information in multiple structures.  Applications usually do not
explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft input that
causes inconsistencies~\cite{evaders6}.  Furthermore, security
vulnerabilities can occur when an application assumes two repetitions
of the same data to be consistent, such as allocating a buffer based on
the value of one size field and copying into that buffer based on the
value of another~\cite{python-bug:20078}.

Finally, real-world data formats, for example PNG or PDF, are hard to
represent with existing parser generators.  Those parsers cannot directly
deal with length or checksum fields directly, so the programmer has to
either write potentially unsafe code to deal with such features, or build
contrived grammar constructs, such as introducing one grammar rule for
each possible value of a length field.  Offset fields, which specify
the position at which some data structure is located, usually require
the programmer to manipulate a parser's internal state to re-position
its input.  More complicated transformations, such as handling compressed
data, cannot be represented at all.

This paper presents the design and implementation of Nail, a parser generator that greatly reduces
the programmer effort required to use grammars. Nail addresses the above four challenges with
several key ideas, as follows.

@@

Second, this well-defined internal representation allows Nail to
maintain a {\em semantic bijection} between data formats and their
internal object model.  As a result, this enables Nail to not just
parse input but also generate output from the internal representation,
without requiring the programmer to write additional code.

>>>>

Code that handles untrusted inputs, such as processing network
data or parsing a file, is error-prone and is often exploited by
attackers.
%   This is in part because attackers have precise control
% over the inputs to that code, and can craft inputs that trigger
% subtle corner cases in input processing.
For example, the libpng image decompression library had 24 remotely
exploitable vulnerabilities from 2007 to 2013~\cite{cvedetails:libpng},
Adobe's PDF and Flash viewers have been notoriously plagued by input
processing vulnerabilities, and even the zlib compression library had
input processing vulnerabilities in the past~\cite{cvedetails:zlib}.
With a memory-unsafe language like C, mistakes in input processing
code can lead to memory errors like buffer overflows, and even with
a memory-safe language like Java, inconsistencies between different
parsers can lead to security issues~\cite{saurik-masterkey}.

A promising approach to avoid these problems is to specify
a precise grammar for the input data format, and to use a parser
generator, such as {\tt lex} and {\tt yacc}, to synthesize the input
processing code.  Developers that use a parser generator do not need
to write error-prone input processing code on their own, and as long
as the parser generator is bug-free, the application will be safe from
input processing vulnerabilities. Grammars can also be re-used between
applications, further reducing effort and eliminating inconsistencies.

@@

Second,  applications often need to produce output in the same
format as their input---for example, applications might both
read and write files, or both receive and send network packets.
Most parser generators focus on just parsing an input, rather
than producing an output, thus requiring the programmer to manually
construct outputs, which is work-intensive and leads to
more code that could contain errors.
Some parser generators, such as
Boost.Spirit~\cite{boost-spirit},
allow reusing the grammar for generating output from the internal
representation.  However, those generators require yet another set of
semantic actions to be written, transforming the internal representation
back into an AST\@.

Third, many data formats contain redundancies, such as repeating
information in multiple structures.  Applications usually do not
explicitly check for consistency, and if different applications use
different instances of the same value, an attacker can craft an input that
causes inconsistencies~\cite{evaders6}.  Furthermore, security
vulnerabilities can occur when an application assumes two repetitions
of the same data to be consistent, such as allocating a buffer based on
the value of one size field and copying into that buffer based on the
value of another~\cite{python-bug:20078}.

Finally, real-world data formats, such as PNG or PDF, are hard to
represent with existing parser generators.  Those parsers cannot directly
deal with length or checksum fields, so the programmer has to
either write potentially unsafe code to deal with such features, or build
contrived grammar constructs, such as introducing one grammar rule for
each possible value of a length field.  Offset fields, which specify
the position at which some data structure is located, usually require
the programmer to manipulate a parser's internal state to re-position
its input.  More complicated transformations, such as handling compressed
data, cannot be represented at all.

This paper presents the design and implementation of Nail, a parser generator that greatly reduces
the programmer effort required to use grammars. Nail addresses the above challenges with
several key ideas, as follows.

@@

Second, this well-defined internal representation allows Nail to
establish a {\em semantic bijection} between data formats and their
internal object model.  As a result, this enables Nail to not just
parse input but also generate output from the internal representation,
without requiring the programmer to write additional code.
